## Analysis for "families with kids waste more"-hypothesis 
## + Identify sample size (mean varies <5%)

```{r}
#0. you might have to install missing packages, comment in the following 2 lines to do so
#install.packages("RColorBrewer")
#install.packages("tidyverse")
library(RColorBrewer)
library(tidyverse)

library(dplyr)
library(purrr)
source("analysis/util.R")



PATH <- "E:/UF/archive/paper_01/data/"

#4. optional: change the design settings: 
save_plots = TRUE
PALETTE = "Spectral"
WIDTH = 1200
HEIGHT = 900
OUTPUT_FOLDER = "plots/"
```

```{r}
#setup dataset, adding std, mean, lower, upper bounds
filterIsServingBased = TRUE 
df_waste_6 <- setup_dataset(PATH,"6_kids/",filterIsServingBased)
df_waste_4 <- setup_dataset(PATH,"4_kids/",filterIsServingBased)
df_waste_2 <- setup_dataset(PATH,"2_kids/",filterIsServingBased)
df_waste_0 <- setup_dataset(PATH,"0_kids/",filterIsServingBased)

```

```{r}
#plot absolute min_max and average
plot = NULL
start = 10
plot <- plot_min_max(df_waste_6,"6 kids",plot,start)
plot <- plot_min_max(df_waste_4,"4 kids",plot,start)
plot <- plot_min_max(df_waste_2,"2 kids",plot,start)
plot <- plot_min_max(df_waste_0,"no kids",plot,start)

plot
```



```{r}
#colors <- brewer.pal(11, PALETTE)[1:4]  # Select the first 4 colors from the Spectral palette


```

```{r}
#TODO color settings
source("analysis/util.R")
#calc confidence interval, plot 
df_waste_0_ci <- calculate_confidence_interval(df_waste_0)
df_waste_2_ci <- calculate_confidence_interval(df_waste_2)
df_waste_4_ci <- calculate_confidence_interval(df_waste_4)
df_waste_6_ci <- calculate_confidence_interval(df_waste_6)

plot2 = NULL

plot2 <- plot_min_max(df_waste_6_ci,"6 children",plot2)
plot2 <- plot_min_max(df_waste_4_ci,"4 children",plot2)
plot2 <- plot_min_max(df_waste_2_ci,"2 children",plot2)
plot2 <- plot_min_max(df_waste_0_ci,"no children",plot2)

plot2

```

```{r}

WIDTH = 10  
HEIGHT = 5
ggsave(plot=plot2, width=WIDTH,height=HEIGHT,filename = "kids_hypo_confidence_interval.png",path = "img", dpi=1200)

```

## Average Food waste per day / per household

```{r}
# Calculate the mean of the 'Mean' column for rows 10 to 100
mean_value_0 <- mean(df_waste_0$Mean[10:100])#/2
mean_value_2 <- mean(df_waste_2$Mean[10:100])#/4
mean_value_4 <- mean(df_waste_4$Mean[10:100])#/6
mean_value_6 <- mean(df_waste_6$Mean[10:100])#/8

mean_vals <- (c(mean_value_0, mean_value_2, mean_value_4, mean_value_6))
mean_vals

```

## Average Food waste per day / per person

```{r}
# Calculate the mean of the 'Mean' column for rows 10 to 100
mean_value_0 <- mean(df_waste_0$Mean[10:100])/2
mean_value_2 <- mean(df_waste_2$Mean[10:100])/4
mean_value_4 <- mean(df_waste_4$Mean[10:100])/6
mean_value_6 <- mean(df_waste_6$Mean[10:100])/8

mean_vals <- (c(mean_value_0, mean_value_2, mean_value_4, mean_value_6))
mean_vals
```

```{r}

columns_to_remove <- c("Day.Wasted", "Mean", "Lower","Upper","Std")
data <- df_waste_6
data <- data[, !names(data) %in% columns_to_remove]
sums <- colSums(data[10:100])/(90*8) #we are averaging over the 90 days and then calc value pp


print(sd(sums))
```

```{r}
length(sums)
```

## Std for food waste per day/per person
```{r}
# Calculate the mean of the 'Mean' column for rows 10 to 100
std_value_0 <- sd(df_waste_0$Mean[10:100])/2
std_value_0

std_value_2 <- sd(df_waste_2$Mean[10:100])/4
std_value_2

std_value_4 <- sd(df_waste_4$Mean[10:100])/6
std_value_4

std_value_6 <- sd(df_waste_6$Mean[10:100])/8
std_value_6

```

## Std for food waste per day/per person
```{r}
# Calculate the mean of the 'Mean' column for rows 10 to 100
std_value_0 <- mean(df_waste_0$Std[10:100])/2
std_value_0

std_value_2 <- mean(df_waste_2$Std[10:100])/4
std_value_2

std_value_4 <- mean(df_waste_4$Std[10:100])/6
std_value_4

std_value_6 <- mean(df_waste_6$Std[10:100])/8
std_value_6

```

# PLOTS FOR PAPER 



```{r}


setup_dataset_all_files<- function(PATH,foldername) {
  # Get the list of subfolder paths
  subfolder_paths <- list.files(paste0(PATH,foldername), full.names = TRUE)
  # Get the list of unique file names across all subfolders
  unique_file_names <- subfolder_paths %>%
    map(~ list.files(., pattern = "\\.csv$", full.names = TRUE)) %>%
    unlist() %>%
    basename() %>%
    unique()
  # Create individual data frames for each unique file name
  df <- unique_file_names %>%
    set_names() %>%
    map(~ {
      file_name_with_slash <- paste0("/",.x)  # Add "/" before unique filename
      # Print the pattern being used
      file_paths <- subfolder_paths %>%
        map(~ {
          files <- (paste0(.x, file_name_with_slash))
        }) %>%
        unlist() %>%
        discard(~ length(.x) == 0)
      map_dfr(file_paths, read.csv,.id = "Run")
    })
  names(df) <- str_remove(names(df), "....$")

  #group by house an run 
  df[["wasted"]]$Group <- paste(df[["wasted"]]$Run,df[["wasted"]]$House,sep="_")
  
  return(df)
}



```

## Plot food waste for specific groups
```{r}
df_6_no_analysis <- setup_dataset_all_files(PATH,"6_kids/")
df_waste_6_no_analysis <- df_6_no_analysis[["wasted"]]
df_config_6_no_analysis <- df_6_no_analysis[["config"]]
df_4_no_analysis <- setup_dataset_all_files(PATH,"4_kids/")
df_waste_4_no_analysis <- df_4_no_analysis[["wasted"]]
df_config_4_no_analysis <- df_4_no_analysis[["config"]]
df_2_no_analysis <- setup_dataset_all_files(PATH,"2_kids/")
df_waste_2_no_analysis <- df_2_no_analysis[["wasted"]]
df_config_2_no_analysis <- df_2_no_analysis[["config"]]
df_0_no_analysis <- setup_dataset_all_files(PATH,"0_kids/")
df_waste_0_no_analysis <- df_0_no_analysis[["wasted"]]
df_config_0_no_analysis <- df_0_no_analysis[["config"]]
```

```{r}

df_waste_all <- rbind(df_waste_0_no_analysis,df_waste_2_no_analysis,df_waste_4_no_analysis, df_waste_6_no_analysis)
df_config_all <- rbind(df_config_0_no_analysis,df_config_2_no_analysis,df_config_4_no_analysis, df_config_6_no_analysis)


# Extract the subset of "Run" and "House" where "IsServingBased" is true in config
serving_runs_houses <- df_config_all$Run[df_config_all$IsServingBased == "True"]
serving_runs_houses <- unique(serving_runs_houses) # Ensure uniqueness
serving_houses <- df_config_all$House[df_config_all$IsServingBased == "True"]
serving_houses <- unique(serving_houses) # Ensure uniquenes
# Filter the wasted dataframe based on serving runs and houses
df_waste_all <- df_waste_all[df_waste_all$Run %in% serving_runs_houses & df_waste_all$House %in% serving_houses, ]
```


```{r}
```

  
```{r}
  #calc food types portions
  #sum_by_type <- aggregate(Kg ~ Type + House, data = df_waste, FUN = sum)
  sum_by_type <- aggregate(Kg ~ Type , data = df_waste_all, FUN = sum)
  sum_by_type$Percentage <- sum_by_type$Kg / sum(sum_by_type$Kg)
  sum_by_type

  sum(sum_by_type$Kg)
```

```{r}


df_waste_all <- rbind(df_waste_0_no_analysis,df_waste_2_no_analysis,df_waste_4_no_analysis, df_waste_6_no_analysis)
df_config_all <- rbind(df_config_0_no_analysis,df_config_2_no_analysis,df_config_4_no_analysis, df_config_6_no_analysis)



filterIsServingBased = TRUE

number_of_people = 2
current_waste_data <- df_waste_all
current_config_data <- df_config_all


if(filterIsServingBased)
{
  # Extract the subset of "Run" and "House" where "IsServingBased" is true in config
  serving_runs_houses <- current_config_data$Run[current_config_data$IsServingBased == "True"]
  serving_runs_houses <- unique(serving_runs_houses) # Ensure uniqueness
  serving_houses <- current_config_data$House[current_config_data$IsServingBased == "True"]
  serving_houses <- unique(serving_houses) # Ensure uniquenes
  # Filter the wasted dataframe based on serving runs and houses

  current_waste_data <- current_waste_data[current_waste_data$Run %in% serving_runs_houses & current_waste_data$House %in% serving_houses, ]
} 
#number of households 
number_of_households <- length(unique(current_waste_data$Group)) 
number_of_households

df_waste_all <- current_waste_data
# Create an empty dataframe
wasted_weight <- data.frame(
  Meat =  sum(df_waste_all$Meat...Fish) * 0.09,
  Veggie =  sum(df_waste_all$Fruits...Vegetables) * 0.116,
  Store =  sum(df_waste_all$Store.Prepared.Items) * 0.095,
  Dry =  sum(df_waste_all$Dry.Foods...Baked.Goods) * 0.065,
  Dairy =  sum(df_waste_all$Dairy...Eggs) * 0.109,
  Snacks =  sum(df_waste_all$Snacks..Condiments..Liquids..Oils..Grease....Other) * 0.095
)

wasted_weight$Total <- wasted_weight$Meat + wasted_weight$Veggie + wasted_weight$Store + wasted_weight$Dry + wasted_weight$Dairy + wasted_weight$Snacks


wasted_weight$Meat_Percentage <- (wasted_weight$Meat  / wasted_weight$Total) 
wasted_weight$Veggie_Percentage <- (wasted_weight$Veggie  / wasted_weight$Total) 
wasted_weight$Store_Percentage <- (wasted_weight$Store  / wasted_weight$Total) 
wasted_weight$Dry_Percentage <- (wasted_weight$Dry  / wasted_weight$Total) 
wasted_weight$Dairy_Percentage <- (wasted_weight$Dairy  / wasted_weight$Total) 
wasted_weight$Snacks_Percentage <- (wasted_weight$Snacks  / wasted_weight$Total) 


wasted_weight$Meat_HH <- (wasted_weight$Meat  / (number_of_households*100)) 
wasted_weight$Veggie_HH <- (wasted_weight$Veggie  / (number_of_households*100)) 
wasted_weight$Store_HH <- (wasted_weight$Store  / (number_of_households*100)) 
wasted_weight$Dry_HH <- (wasted_weight$Dry  / (number_of_households*100)) 
wasted_weight$Dairy_HH <- (wasted_weight$Dairy  / (number_of_households*100)) 
wasted_weight$Snacks_HH <- (wasted_weight$Snacks  / (number_of_households*100)) 
wasted_weight$Total_HH <- wasted_weight$Meat_HH + wasted_weight$Veggie_HH + wasted_weight$Store_HH + wasted_weight$Dry_HH + wasted_weight$Dairy_HH + wasted_weight$Snacks_HH


wasted_weight$Meat_PP <- (wasted_weight$Meat  / (number_of_households*100*number_of_people)) 
wasted_weight$Veggie_PP <- (wasted_weight$Veggie  / (number_of_households*100*number_of_people)) 
wasted_weight$Store_PP <- (wasted_weight$Store  / (number_of_households*100*number_of_people)) 
wasted_weight$Dry_PP <- (wasted_weight$Dry  / (number_of_households*100*number_of_people)) 
wasted_weight$Dairy_PP <- (wasted_weight$Dairy  / (number_of_households*100*number_of_people)) 
wasted_weight$Snacks_PP <- (wasted_weight$Snacks  / (number_of_households*100*number_of_people)) 
wasted_weight$Total_PP <- wasted_weight$Meat_PP + wasted_weight$Veggie_PP + wasted_weight$Store_PP + wasted_weight$Dry_PP + wasted_weight$Dairy_PP + wasted_weight$Snacks_PP


wasted_weight


```

## Waste per kids amount group / only servingbased

```{r}
```

```{r}

```

## Sum by Waste type
```{r}
  #calc food types portions
  #sum_by_type <- aggregate(Kg ~ Type + House, data = df_waste, FUN = sum)
  sum_by_status <- aggregate(Kg ~ Status , data = df_waste_all, FUN = sum)
  sum_by_status$Percentage <- sum_by_status$Kg / sum(sum_by_status$Kg)
  sum_by_status

  sum(sum_by_status$Kg)
```


```{r}
### ensure you selected serving based ones

df_waste_0_no_analysis <- subset(df_waste_0_no_analysis, Run == 1 & House == 1)
df_config_0_no_analysis <- subset(df_config_0_no_analysis, Run == 1 & House == 1)


df_waste_2_no_analysis <- subset(df_waste_2_no_analysis, Run == 1 & House == 2)
df_config_2_no_analysis <- subset(df_config_2_no_analysis, Run == 1 & House == 2)


df_waste_4_no_analysis <- subset(df_waste_4_no_analysis, Run == 1 & House == 3)
df_config_4_no_analysis <- subset(df_config_4_no_analysis, Run == 1 & House == 3)


df_waste_6_no_analysis <- subset(df_waste_6_no_analysis, Run == 1 & House == 4)
df_config_6_no_analysis <- subset(df_config_6_no_analysis, Run == 1 & House == 4)

df_waste_no_analysis <- rbind(df_waste_0_no_analysis,df_waste_2_no_analysis,df_waste_4_no_analysis, df_waste_6_no_analysis)
df_config_no_analysis <- rbind(df_config_0_no_analysis,df_config_2_no_analysis,df_config_4_no_analysis, df_config_6_no_analysis)

```
```{r}
```

```{r}
PALETTE = "Spectral"
 #Create named vector with labels for each house
label_house_num_people <- as.vector(
  with(
    df_config_no_analysis,
    sprintf("House: %s, Adults: %i, Children: %i Concern: %.2f", House, Adults, Children, LvlOfConcern)
  )
)



# Plotting
wasted_by_household_fw_type <- ggplot(data = df_waste_no_analysis) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Status, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, df_config_no_analysis$House)))) +
  ylab("Kg") + 
  xlab("Day") + 
  scale_fill_brewer(palette = PALETTE) +
  theme_minimal()

# Display the plot
wasted_by_household_fw_type

```



## Check if average still changes 
```{r}

#default #10hh per run 
df_waste_0_2 <- setup_dataset(PATH,"0_kids_40/") #20hh per run 
df_waste_0_3 <- setup_dataset(PATH,"0_kids_50hh/") #50hh per run
plot3 <- NULL

plot3 <- plot_lower_upper(df_waste_0,"",plot3)
plot3 <- plot_lower_upper(df_waste_0_2,"",plot3)
plot3 <- plot_lower_upper(df_waste_0_3,"",plot3)

plot3
```

```{r}

Mean <- rowMeans(df_waste_0[, -1])
mean(Mean)
Mean2 <- rowMeans(df_waste_0_2[, -1])
mean(Mean2)

Mean3 <- rowMeans(df_waste_0_3[, -1])
mean(Mean3)


mean(Mean2)/mean(Mean)
mean(Mean3)/mean(Mean2)


```

