```{r}
#0. you might have to install missing packages, comment in the following 2 lines to do so
#install.packages("RColorBrewer")
#install.packages("tidyverse")
library(RColorBrewer)
library(tidyverse)

```

## Importing the current dataset + setup 

```{r}
##if you want to run the code, you have to change the path to your folder
#you can run main.py with you configuration, which will create a new folder with a timestamp
#in the datafolder
#1. add the path to the data folder in PATH
#2. add the name of the folder you want to analyse e.g.  "2024-04-10at09-45/"
PATH = "E:/UF/ifwaste/data/"
INPUT_FOLDER = "run_0/"

#3. if you want to save all generated plots, set save_plots to TRUE, otherwise set it to FALSE
save_plots = TRUE

#4. optional: change the design settings: 
PALETTE = "Spectral"
WIDTH = 1200
HEIGHT = 900


OUTPUT_FOLDER = "plots/"

hh_foodwasted <- read.csv(paste0(PATH, INPUT_FOLDER, "/wasted.csv"))
hh_foodeaten <- read.csv(paste0(PATH, INPUT_FOLDER, "/eaten.csv"))
hh_foodremaining <- read.csv(paste0(PATH, INPUT_FOLDER, "/still_have.csv"))
hh_foodbought <- read.csv(paste0(PATH, INPUT_FOLDER, "/bought.csv"))
hh_daily <- read.csv(paste0(PATH, INPUT_FOLDER, "/daily.csv"))
hh_config <- read.csv(paste0(PATH, INPUT_FOLDER, "/config.csv"))

```

Label generator for plots
```{r}

# Create named vector with labels for each house
label_house_num_people <- as.vector(
  with(
    hh_config,
    sprintf("House: %s, Adults: %i, Children: %i Concern: %.2f", House, Adults, Children, LvlOfConcern)
  )
)

house_labels <- function(x) paste0("House ", x)
```

## Biomass check: 
### 1. in kg
```{r}

remaining <- aggregate(Kg ~ Day, data = hh_foodremaining, FUN = sum)
remaining <- tail(remaining, n = 1)$Kg
tail(hh_foodremaining, n = 1)
print("bought:")
sum(hh_foodbought$Kg)
print("wasted:")
sum(hh_foodwasted$Kg)
print("eaten:")
sum(hh_foodeaten$Kg)
print("remained:")
remaining

print("total:")

total <- sum(hh_foodwasted$Kg) + sum(hh_foodeaten$Kg) + remaining
total 

print("difference b-t:")
missing <- sum(hh_foodbought$Kg) - total
missing 
```
### 2. in servings
```{r}

remaining <- aggregate(Servings ~ Day, data = hh_foodremaining, FUN = sum)
remaining <- tail(remaining, n = 1)$Servings
tail(hh_foodremaining, n = 1)
print("bought:")
sum(hh_foodbought$Servings)
print("wasted:")
sum(hh_foodwasted$Servings)
print("eaten:")
sum(hh_foodeaten$Servings)
print("remained:")
remaining

print("total:")

total <- sum(hh_foodwasted$Servings) + sum(hh_foodeaten$Servings) + remaining
total 

print("difference b-t:")
missing <- sum(hh_foodbought$Servings) - total
missing 
```
---------------------------------------------------------
## Analysis of food purchase
### 1. Purchased food by household divided by type, in servings

```{r}

# Plotting
purchased_by_household <- ggplot(data = hh_foodbought) +
  geom_bar(mapping = aes(x = Day.Bought, fill = Type, weight = Servings), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in servings") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
purchased_by_household

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "bought_hh_servings.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(purchased_by_household)
   dev.off()
}


```


---------------------------------------------------------
## Analysis of food waste 
### 1. Food waste by household by kg

```{r}

# Plotting
wasted_by_household <- ggplot(data = hh_foodwasted) +
  stat_summary(aes(x = Day.Wasted, y = Kg), fun="sum", geom="point") +
  stat_summary(aes(x = Day.Wasted, y = Kg), fun="sum", geom="line") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE) 

wasted_by_household

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household)
   dev.off()
}

```

## Food waste per household and food type
### 2. Food waste by household divided by type in kg
```{r}
# Plotting
wasted_by_household_kg <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE)

# Display the plot
wasted_by_household_kg

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_  ype.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_kg)
   dev.off()
}

```

### 3. Food waste by household divided by food waste type in kg

```{r}

# Plotting
wasted_by_household_fw_type <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Status, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_fw_type

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_fwtype_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_fw_type)
   dev.off()
}


```

### 4. Food waste by household divided by food type in USD
```{r}
# Plotting
wasted_by_household_price <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Price), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in USD") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_price

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_fwtype_usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_price)
   dev.off()
}
```
### 5. Food waste type in kg

```{r}
wasted_by_type = ggplot(data=hh_foodwasted) +
    geom_bar(mapping = aes(x=Type, fill=House, weight=Kg), position="stack") +
    ylab("in Kg") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    scale_fill_brewer(palette = PALETTE) 
wasted_by_type


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_type_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_type)
   dev.off()
}
```

### 6. Total food wasted in Kg by food waste type: 

```{r}
sum_by_status <- aggregate(Kg ~ Status, data=hh_foodwasted, FUN = sum)
print(sum_by_status)
```

## Nutrients and Behavior

### 1. Missing servings and kcals per household 

```{r}
scaleFactor <- max(hh_daily$Serving)/ max(hh_daily$Kcal)

consumption_serv_kcal = ggplot(data=hh_daily) + 
    geom_line(mapping = aes(x=Day, y=Servings, group=House), color="#216073") +
    geom_line(mapping = aes(x=Day, y=Kcal*scaleFactor, group=House), color="#93144b") + 
    scale_y_continuous(name="Missing servings", sec.axis=sec_axis(~./scaleFactor, name="Missing kcal")) +
    facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
    theme(
        axis.title.y.left=element_text(color="#216073"),
        axis.text.y.left=element_text(color="#216073"),
        axis.title.y.right=element_text(color="#93144b"),
        axis.text.y.right=element_text(color="#93144b")
    )  +
  scale_fill_brewer(palette = PALETTE) 

consumption_serv_kcal

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_serv_kcal.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumption_serv_kcal)
   dev.off()
}
```

### 2. Eaten servings in relation to what is required
```{r}
consumed_by_household = ggplot(data=NULL) +
    geom_bar(data=hh_foodeaten, mapping = aes(x=Day.Eaten, fill=Type, weight=Servings), position="stack") +
    geom_hline(data=hh_config, mapping= aes(yintercept=RequiredServings, linetype="Threshold")) + 
    facet_wrap(~House, labeller = as_labeller(house_labels)) +
    ylab("in servings") + 
    scale_fill_brewer(palette = PALETTE) + 
    scale_linetype_manual(values = c("dashed"), labels = c("Required servings"), name = NULL) 

consumed_by_household


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_serv_reqserv.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumed_by_household)
   dev.off()
}

``` 

### 3. Eating kcal in relation to what is required
```{r}
consumed_by_household = ggplot(data=NULL) +
    geom_bar(data=hh_foodeaten, mapping = aes(x=Day.Eaten, fill=Type, weight=Kcal), position="stack") +
    geom_hline(data=hh_config, mapping= aes(yintercept=RequiredKcal, linetype="Threshold")) + 
    facet_wrap(~House, labeller = as_labeller(house_labels)) +
    ylab("in kcal") + 
    scale_fill_brewer(palette = PALETTE) + 
    scale_linetype_manual(values = c("dashed"), labels = c("Required servings"), name = NULL) 

consumed_by_household


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_kcal_reqkcal.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumed_by_household)
   dev.off()
}

``` 

### 3. Meal preparation choice distribution 
```{r}

# Reshape data to long format
hh_daily_long <- pivot_longer(hh_daily, cols = c("EEF","Cooked", "QuickCook","AteLeftOvers"), names_to = "Action", values_to = "Value")


label_house_concern <- as.vector(
  with(
    hh_config,
    sprintf("House: %s Concern: %.2f", House, LvlOfConcern)  
)) 

# Create the plot
plot <- ggplot(data = hh_daily_long, aes(x = Day, y = Value, fill = Action)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(rows = vars(Action), cols = vars(House), 
      scales = "free",
      labeller = labeller(House = as_labeller(setNames(label_house_concern, hh_config$House)))) +
  labs(x = "Days",
       y = "1 = action performed, 0 = not performed") +
  scale_fill_brewer(palette = PALETTE) 

plot

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "meal_prep_behavior.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(plot)
   dev.off()
}

```


## Budget
### 1. Budget per household

```{r}

budget = ggplot(data=hh_daily) + 
    geom_line(mapping = aes(x=Day, y=Budget, group=House)) +
    facet_wrap(~House, nrow=2,labeller = as_labeller(house_labels))

budget


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(budget)
   dev.off()
}
```

### 2. Money spend on food vs. money wasted due to food waste

```{r}

sum_by_status <- aggregate(Price ~ Type, data=hh_foodbought, FUN = sum)

# Plotting
bought_vs_wasted <- ggplot(data =NULL) +
  geom_bar(data=hh_foodbought, mapping = aes(x = Day.Bought, weight = Price, fill="Bought")) +
  geom_bar(data=hh_foodwasted, mapping = aes(x = Day.Wasted, weight = Price, fill="Wasted")) +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in USD") + 
  scale_fill_brewer(palette = PALETTE)
  #scale_fill_manual(values = c( "Bought" = "#6b6668", "Wasted" = "#a8325e")) 

# Display the plot
bought_vs_wasted


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "bought_waste_usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(bought_vs_wasted)
   dev.off()
}

```


```{r}

#spendings_buying <- aggregate(Price ~ Type, data=hh_foodbought, FUN = sum)
#spendings_wasting <- aggregate(Price ~ Type, data=hh_foodwasted, FUN = sum)
#difference <- spendings_buying - spendings_wasting
#time <- hh_foodbought$Day.Bought
# Plotting
#wasted_by_household <- ggplot(data = NULL) +
#    geom_col(data = difference)
#  geom_bar(data=hh_foodbought, mapping = aes(x = Day.Bought, weight = Price, fill="Bought")) +
#  geom_bar(data=hh_foodwasted, mapping = aes(x = Day.Wasted, weight = Price, fill="Wasted")) +
#  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
#  ylab("in USD") + 
#  scale_fill_brewer(palette = PALETTE) + 
#  scale_fill_manual(values = c( "Bought" = "#6b6668", "Wasted" = "#a8325e")) +
#  #ggtitle("Spendings for groceries vs lost value due to waste")

# Display the plot
#print(wasted_by_household)

```
```{r}                  

### TODO: split by household and sum 


#spendings_buying <- aggregate(Price ~ Day.Bought , data=hh_foodbought, FUN = sum)
#spendings_wasting <- aggregate(Price ~ Day.Wasted , data=hh_foodwasted, FUN = sum)

#difference <- spendings_buying - spendings_wasting
#time <- unique(hh_foodbought$Day.Bought)

#print(time)
#print(difference)

#df <- data.frame( 
#    "difference" = c(difference),
#    "time" = c(time)
#)

#df$direction <- ifelse(df$difference >= 0, "Positive", "Negative")

# Plot the differences
#ggplot(df, aes(x = time, y = difference, fill = direction)) +
#  geom_bar(stat = "identity", position = "identity") +
#  scale_fill_manual(values = c("Positive" = "blue", "Negative" = "red")) +
#  labs(x = "Category", y = "Difference", title = "Difference between Values") +
#  theme_minimal()

```
