
```{r}
source("analysis/util.R")

test()

```
```{r}
#0. you might have to install missing packages, comment in the following 2 lines to do so
#install.packages("RColorBrewer")
#install.packages("tidyverse")
library(RColorBrewer)
library(tidyverse)

```

## Importing the current dataset + setup 

```{r}
##if you want to run the code, you have to change the path to your folder
#you can run main.py with you configuration, which will create a new folder with a timestamp
#in the datafolder
#1. add the path to the data folder in PATH
#2. add the name of the folder you want to analyse e.g.  "2024-04-10at09-45/"
PATH = "E:/UF/ifwaste/archive/01_paper/data/2_kids/"


#INPUT_FOLDER = "run_0/"

#3. if you want to save all generated plots, set save_plots to TRUE, otherwise set it to FALSE
save_plots = FALSE

#4. optional: change the design settings: 
PALETTE = "Spectral"
WIDTH = 1200
HEIGHT = 900


OUTPUT_FOLDER = "plots/"

hh_foodwasted <- read.csv(paste0(PATH, INPUT_FOLDER, "/wasted.csv"))
hh_foodeaten <- read.csv(paste0(PATH, INPUT_FOLDER, "/eaten.csv"))
hh_foodremaining <- read.csv(paste0(PATH, INPUT_FOLDER, "/still_have.csv"))
hh_foodbought <- read.csv(paste0(PATH, INPUT_FOLDER, "/bought.csv"))
hh_daily <- read.csv(paste0(PATH, INPUT_FOLDER, "/daily.csv"))
hh_config <- read.csv(paste0(PATH, INPUT_FOLDER, "/config.csv"))

```

Label generator for plots
```{r}

# Create named vector with labels for each house
label_house_num_people <- as.vector(
  with(
    hh_config,
    sprintf("House: %s, Adults: %i, Children: %i Concern: %.2f", House, Adults, Children, LvlOfConcern)
  )
)

house_labels <- function(x) paste0("House ", x)
```

## Biomass check: 
### 1. in kg
```{r}

remaining <- aggregate(Kg ~ Day, data = hh_foodremaining, FUN = sum)
remaining <- tail(remaining, n = 1)$Kg
tail(hh_foodremaining, n = 1)
print("bought:")
sum(hh_foodbought$Kg)
print("wasted:")
sum(hh_foodwasted$Kg)
print("eaten:")
sum(hh_foodeaten$Kg)
print("remained:")
remaining

print("total:")

total <- sum(hh_foodwasted$Kg) + sum(hh_foodeaten$Kg) + remaining
total 

print("difference b-t:")
missing <- sum(hh_foodbought$Kg) - total
missing 
```
### 2. in servings
```{r}

remaining <- aggregate(Servings ~ Day, data = hh_foodremaining, FUN = sum)
remaining <- tail(remaining, n = 1)$Servings
tail(hh_foodremaining, n = 1)
print("bought:")
sum(hh_foodbought$Servings)
print("wasted:")
sum(hh_foodwasted$Servings)
print("eaten:")
sum(hh_foodeaten$Servings)
print("remained:")
remaining

print("total:")

total <- sum(hh_foodwasted$Servings) + sum(hh_foodeaten$Servings) + remaining
total 

print("difference b-t:")
missing <- sum(hh_foodbought$Servings) - total
missing 
```
---------------------------------------------------------
## Analysis of food purchase
### 1. Purchased food by household divided by type, in servings

```{r}

# Plotting
purchased_by_household <- ggplot(data = hh_foodbought) +
  geom_bar(mapping = aes(x = Day.Bought, fill = Type, weight = Servings), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in servings") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
purchased_by_household

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "bought_hh_servings.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(purchased_by_household)
   dev.off()
}


```


---------------------------------------------------------
## Analysis of food waste 
### 1. Food waste by household by kg

```{r}

# Plotting
wasted_by_household <- ggplot(data = hh_foodwasted) +
  stat_summary(aes(x = Day.Wasted, y = Kg), fun="sum", geom="point") +
  stat_summary(aes(x = Day.Wasted, y = Kg), fun="sum", geom="line") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE) 

wasted_by_household

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household)
   dev.off()
}

```

## Food waste per household and food type
### 2. Food waste by household divided by type in kg
```{r}
# Plotting
wasted_by_household_kg <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE)

# Display the plot
wasted_by_household_kg

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_type.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_kg)
   dev.off()
}

```

### 3. Food waste by household divided by food waste type in kg

```{r}

# Plotting
wasted_by_household_fw_type <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Status, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_fw_type

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_fwtype_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_fw_type)
   dev.off()
}


```

```{r}
wasted_kg <- aggregate(Kg ~ Day.Wasted+House, data = hh_foodwasted, FUN = sum)
sum_kg_by_house <- wasted_kg %>%
  group_by(House) %>%
  summarise(Total_Kg = sum(Kg)/(40*6))
sum_kg_by_house

#sum(sum_kg_by_house$Total_Kg)

#sum(sum_kg_by_house$Total_Kg[0:5])
#sum(sum_kg_by_house$Total_Kg[5:9])
```

### 4. Food waste by household divided by food type in USD
```{r}
# Plotting
wasted_by_household_price <- ggplot(data = hh_foodwasted) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Price), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in USD") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_price

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_fwtype_usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_price)
   dev.off()
}
```
### 5. Food waste type in kg

```{r}
wasted_by_type = ggplot(data=hh_foodwasted) +
    geom_bar(mapping = aes(x=Type, weight=Kg), position="stack") +
    ylab("in Kg") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  
    #scale_fill_brewer(palette = PALETTE) 
wasted_by_type


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_type_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_type)
   dev.off()
}
```

### 6. Total food wasted in Kg by food waste type: 

```{r}
sum_by_status <- aggregate(Kg ~ Status, data=hh_foodwasted, FUN = sum)
print(sum_by_status)
```

## Still have
```{r}

# Plotting
purchased_by_household <- ggplot(data = hh_foodremaining) +
  geom_bar(mapping = aes(x = Day, fill = Type, weight = Servings), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in servings") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
purchased_by_household

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "bought_hh_servings.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(purchased_by_household)
   dev.off()
}


```

## Nutrients and Behavior

### 1. Missing servings and kcals per household 


## eaten servings
```{r}
eaten <- aggregate(Servings ~ Day.Eaten+House, data = hh_foodeaten, FUN = sum)
eaten

```
## eaten kcal
```{r}
eaten <- aggregate(Kcal ~ Day.Eaten+House, data = hh_foodeaten, FUN = sum)
eaten

```

# food eaten by type in kg
```{r}
# Plotting
wasted_by_household_fw_type <- ggplot(data = hh_foodeaten) +
  geom_bar(mapping = aes(x = Day.Eaten, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in Kg") + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_fw_type

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "waste_hh_fwtype_kg.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(wasted_by_household_fw_type)
   dev.off()
}
```

```{r}
scaleFactor <- max(hh_daily$Serving)/ max(hh_daily$Kcal)

consumption_serv_kcal = ggplot(data=hh_daily) + 
    geom_line(mapping = aes(x=Day, y=Servings, group=House), color="#216073") +
    geom_line(mapping = aes(x=Day, y=Kcal*scaleFactor, group=House), color="#93144b") + 
    scale_y_continuous(name="Missing servings", sec.axis=sec_axis(~./scaleFactor, name="Missing kcal")) +
    facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
    theme(
        axis.title.y.left=element_text(color="#216073"),
        axis.text.y.left=element_text(color="#216073"),
        axis.title.y.right=element_text(color="#93144b"),
        axis.text.y.right=element_text(color="#93144b")
    )  +
  scale_fill_brewer(palette = PALETTE) 

consumption_serv_kcal

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_serv_kcal.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumption_serv_kcal)
   dev.off()
}
```

### 2. Eaten servings in relation to what is required
```{r}
consumed_by_household = ggplot(data=NULL) +
    geom_bar(data=hh_foodeaten, mapping = aes(x=Day.Eaten, fill=Type, weight=Servings), position="stack") +
    geom_hline(data=hh_config, mapping= aes(yintercept=RequiredServings, linetype="Threshold")) + 
    facet_wrap(~House, labeller = as_labeller(house_labels)) +
    ylab("in servings") + 
    scale_fill_brewer(palette = PALETTE) + 
    scale_linetype_manual(values = c("dashed"), labels = c("Required servings"), name = NULL) 

consumed_by_household


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_serv_reqserv.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumed_by_household)
   dev.off()
}

``` 

### 3. Eating kcal in relation to what is required
```{r}
consumed_by_household = ggplot(data=NULL) +
    geom_bar(data=hh_foodeaten, mapping = aes(x=Day.Eaten, fill=Type, weight=Kcal), position="stack") +
    geom_hline(data=hh_config, mapping= aes(yintercept=RequiredKcal, linetype="Threshold")) + 
    facet_wrap(~House, labeller = as_labeller(house_labels)) +
    ylab("in kcal") + 
    scale_fill_brewer(palette = PALETTE) + 
    scale_linetype_manual(values = c("dashed"), labels = c("Required servings"), name = NULL) 

consumed_by_household


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "eaten_kcal_reqkcal.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(consumed_by_household)
   dev.off()
}

``` 

### 3. Meal preparation choice distribution 
```{r}

# Reshape data to long format
hh_daily_long <- pivot_longer(hh_daily, cols = c("EEF","Cooked", "QuickCook","AteLeftOvers"), names_to = "Action", values_to = "Value")


label_house_concern <- as.vector(
  with(
    hh_config,
    sprintf("House: %s Concern: %.2f", House, LvlOfConcern)  
)) 

# Create the plot
plot <- ggplot(data = hh_daily_long, aes(x = Day, y = Value, fill = Action)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(rows = vars(Action), cols = vars(House), 
      scales = "free",
      labeller = labeller(House = as_labeller(setNames(label_house_concern, hh_config$House)))) +
  labs(x = "Days",
       y = "1 = action performed, 0 = not performed") +
  scale_fill_brewer(palette = PALETTE) 

plot

if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "meal_prep_behavior.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(plot)
   dev.off()
}

```


## Budget
### 1. Budget per household

```{r}

budget = ggplot(data=hh_daily) + 
    geom_line(mapping = aes(x=Day, y=Budget, group=House)) +
    facet_wrap(~House, nrow=2,labeller = as_labeller(house_labels))

budget


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(budget)
   dev.off()
}
```

### 2. Money spend on food vs. money wasted due to food waste

```{r}

sum_by_status <- aggregate(Price ~ Type, data=hh_foodbought, FUN = sum)

# Plotting
bought_vs_wasted <- ggplot(data =NULL) +
  geom_bar(data=hh_foodbought, mapping = aes(x = Day.Bought, weight = Price, fill="Bought")) +
  geom_bar(data=hh_foodwasted, mapping = aes(x = Day.Wasted, weight = Price, fill="Wasted")) +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
  ylab("in USD") + 
  scale_fill_brewer(palette = PALETTE)
  #scale_fill_manual(values = c( "Bought" = "#6b6668", "Wasted" = "#a8325e")) 

# Display the plot
bought_vs_wasted


if (save_plots) {
   png(file = paste0(OUTPUT_FOLDER, "bought_waste_usd.png"), width = WIDTH, height = HEIGHT)  # Adjust width and height as needed
   print(bought_vs_wasted)
   dev.off()
}

```

### 1 Household: waste in kg, usd, kcal

```{r}

subset_hh_foodwasted <- subset(hh_foodwasted, House == 4)

```
```{r}
library(gridExtra)
library(cowplot)

# Plot for Price
plot_price <- ggplot(subset_hh_foodwasted, aes(x = Day.Wasted, fill=Type, y = Price)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "House with 2 adults, 2 children", x = "Day", y = "USD") + 
  scale_fill_brewer(palette = PALETTE) +
  xlim(75,90)

# Plot for Kg
plot_kg <- ggplot(subset_hh_foodwasted, aes(x = Day.Wasted, fill=Type, y = Kg)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "", x = "Day", y = "Kg") + 
  scale_fill_brewer(palette = PALETTE) +
  guides(fill = FALSE) + # Hide legend for this plot
  xlim(75,90)

# Plot for Kcal
plot_kcal <- ggplot(subset_hh_foodwasted, aes(x = Day.Wasted, fill=Type, y = Kcal)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "", x = "Day", y = "Kcal") + 
  scale_fill_brewer(palette = PALETTE) +
  guides(fill = FALSE) + # Hide legend for this plot
  xlim(75,90)

# Combine plots and legend
subplot <- plot_grid(plot_price, plot_kg, plot_kcal, align = "v", ncol = 1, rel_heights = c(1, 1, 1),
                      hjust = -1, # Adjust legend position
                      label_size = 10) + 
  theme(legend.position = "bottom")  # Position legend at the bottom

subplot



```

```{r}
# Create a new plotting window with three subplots arranged vertically
par(mfrow=c(3,1))

# Plot Price
barplot(hh_foodwasted$Day.Wasted, hh_foodwasted$Price, xlab="Day.Wasted", ylab="Price")

# Plot Kg
barplot(hh_foodwasted$Day.Wasted, hh_foodwasted$Kg, xlab="Day.Wasted", ylab="Kg")

# Plot Kcal
barplot(hh_foodwasted$Day.Wasted, hh_foodwasted$Kcal, xlab="Day.Wasted", ylab="Kcal")

```

```{r}

#spendings_buying <- aggregate(Price ~ Type, data=hh_foodbought, FUN = sum)
#spendings_wasting <- aggregate(Price ~ Type, data=hh_foodwasted, FUN = sum)
#difference <- spendings_buying - spendings_wasting
#time <- hh_foodbought$Day.Bought
# Plotting
#wasted_by_household <- ggplot(data = NULL) +
#    geom_col(data = difference)
#  geom_bar(data=hh_foodbought, mapping = aes(x = Day.Bought, weight = Price, fill="Bought")) +
#  geom_bar(data=hh_foodwasted, mapping = aes(x = Day.Wasted, weight = Price, fill="Wasted")) +
#  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, hh_config$House)))) +
#  ylab("in USD") + 
#  scale_fill_brewer(palette = PALETTE) + 
#  scale_fill_manual(values = c( "Bought" = "#6b6668", "Wasted" = "#a8325e")) +
#  #ggtitle("Spendings for groceries vs lost value due to waste")

# Display the plot
#print(wasted_by_household)

```
```{r}                  

### TODO: split by household and sum 


#spendings_buying <- aggregate(Price ~ Day.Bought , data=hh_foodbought, FUN = sum)
#spendings_wasting <- aggregate(Price ~ Day.Wasted , data=hh_foodwasted, FUN = sum)

#difference <- spendings_buying - spendings_wasting
#time <- unique(hh_foodbought$Day.Bought)

#print(time)
#print(difference)

#df <- data.frame( 
#    "difference" = c(difference),
#    "time" = c(time)
#)

#df$direction <- ifelse(df$difference >= 0, "Positive", "Negative")

# Plot the differences
#ggplot(df, aes(x = time, y = difference, fill = direction)) +
#  geom_bar(stat = "identity", position = "identity") +
#  scale_fill_manual(values = c("Positive" = "blue", "Negative" = "red")) +
#  labs(x = "Category", y = "Difference", title = "Difference between Values") +
#  theme_minimal()

```

```{r}


PATH <- "E:/UF/ifwaste/data/"

setup_dataset<- function(PATH,foldername) {
  # Get the list of subfolder paths
  subfolder_paths <- list.files(paste0(PATH,foldername), full.names = TRUE)
  # Get the list of unique file names across all subfolders
  unique_file_names <- subfolder_paths %>%
    map(~ list.files(., pattern = "\\.csv$", full.names = TRUE)) %>%
    unlist() %>%
    basename() %>%
    unique()
  # Create individual data frames for each unique file name
  df <- unique_file_names %>%
    set_names() %>%
    map(~ {
      file_name_with_slash <- paste0("/",.x)  # Add "/" before unique filename
      # Print the pattern being used
      file_paths <- subfolder_paths %>%
        map(~ {
          files <- (paste0(.x, file_name_with_slash))
        }) %>%
        unlist() %>%
        discard(~ length(.x) == 0)
      map_dfr(file_paths, read.csv,.id = "Run")
    })
  names(df) <- str_remove(names(df), "....$")

  #group by house an run 
  df[["wasted"]]$Group <- paste(df[["wasted"]]$Run,df[["wasted"]]$House,sep="_")
  #wasted <- df[["wasted"]] #for now only the wasted subfolder is relevant
  
  return(df)
}



```

# PLOTS FOR PAPER 

## Plot food waste for specific groups
```{r}
df_6 <- setup_dataset(PATH,"6_kids/")
df_waste_6 <- df_6[["wasted"]]
df_config_6 <- df_6[["config"]]
df_4 <- setup_dataset(PATH,"4_kids/")
df_waste_4 <- df_4[["wasted"]]
df_config_4 <- df_4[["config"]]
df_2 <- setup_dataset(PATH,"2_kids/")
df_waste_2 <- df_2[["wasted"]]
df_config_2 <- df_2[["config"]]
df_0 <- setup_dataset(PATH,"0_kids/")
df_waste_0 <- df_0[["wasted"]]
df_config_0 <- df_0[["config"]]
```

```{r}

df_waste_0 <- subset(df_waste_0, Run == 1 & (House == 1 | House == 2))
df_config_0 <- subset(df_config_0, Run == 1 & (House == 1 | House == 2))

df_waste_2 <- subset(df_waste_2, Run == 1 & (House == 3 | House == 4))
df_config_2 <- subset(df_config_2, Run == 1 & (House == 3 | House == 4))


df_waste_4 <- subset(df_waste_4, Run == 1 & (House == 5 | House == 6))
df_config_4 <- subset(df_config_4, Run == 1 & (House == 5 | House == 6))


df_waste_6 <- subset(df_waste_6, Run == 1 & (House == 7 | House == 8))
df_config_6 <- subset(df_config_6, Run == 1 & (House == 7 | House == 8))


df_waste <- rbind(df_waste_0,df_waste_2,df_waste_4, df_waste_6)
df_config <- rbind(df_config_0,df_config_2,df_config_4, df_config_6)

```
```{r}
df_config
```


```{r}

PALETTE = "Spectral"
 #Create named vector with labels for each house
label_house_num_people <- as.vector(
  with(
    df_config,
    sprintf("Adults: %i, Children: %i Concern: %.2f", Adults, Children, LvlOfConcern)
  )
)



# Plotting
wasted_by_household_fw_type <- ggplot(data = df_waste) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, df_config$House)))) +
  ylab("Food wasted (kg)") + 
  xlab("Simulated Days (day)") + 
  #guides(fill=guide_legend(title="Food group"))+
  scale_fill_brewer(palette = PALETTE) +
  theme_minimal() + 
  theme(text = element_text(size = 14), legend.position = "none")

# Display the plot
wasted_by_household_fw_type

```

```{r}

WIDTH = 14
HEIGHT = 7
ggsave(plot=wasted_by_household_fw_type, width=WIDTH,height=HEIGHT,filename = "household_food_waste.png",path = "img", dpi=1200)
```

```{r}

library(ggplot2)

# Plotting without legend
wasted_by_household_fw_type <- ggplot(data = df_waste) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, df_config$House)))) +
  ylab("Food wasted (kg)") + 
  xlab("Simulated Days (day)") + 
  scale_fill_brewer(palette = PALETTE) +
  theme_minimal()

# Display the plot
wasted_by_household_fw_type

```

```{r}
# Plotting
wasted_by_household_kg <- ggplot(data = df_waste) +
  geom_bar(mapping = aes(x = Day.Wasted, fill = Type, weight = Kg), position = "stack") +
  facet_wrap(~House, labeller = labeller(House = as_labeller(setNames(label_house_num_people, df_config$House)))) +
  ylab("Kg") + 
  xlab("Day") + 
  theme_minimal() + 
  scale_fill_brewer(palette = PALETTE) 

# Display the plot
wasted_by_household_kg

```


## Sum by Food type 

```{r}
#calc food types portions
#sum_by_type <- aggregate(Kg ~ Type + House, data = df_waste, FUN = sum)
sum_by_type <- aggregate(Kg ~ Type , data = df_waste, FUN = sum)
sum_by_type$Percentage <- sum_by_type$Kg / sum(sum_by_type$Kg)
sum_by_type

sum(sum_by_type$Kg)
```

## Sum by Waste type
```{r}
#calc food types portions
#sum_by_type <- aggregate(Kg ~ Type + House, data = df_waste, FUN = sum)
sum_by_status <- aggregate(Kg ~ Status , data = df_waste, FUN = sum)
sum_by_status$Percentage <- sum_by_status$Kg / sum(sum_by_status$Kg)
sum_by_status

sum(sum_by_status$Kg)
```

